---
title: "cleaning"
output: html_document
date: "2022-12-15"
bibliography: references.bib
---

```{r, message=FALSE}
library(tidyverse)
library(haven) # read stata data
```

For all data sets we extract the following variables: 

- `trust`
- `id`
- `country`
- `date` (for panel data sets that are on a smaller scale than years)
- `year`
- `scale`

For some data sets, we might export additional variables. 

# General Social Survey US {gss_cross}

```{r, message=FALSE}
# read data
gss_cross <- read_dta("./data/gss_cross/gss7222_r2.dta")
```

## Trust

-   `CONSCI`: "(I am going to name some institutions in this country. As far as the people running these institutions are concerned, would you say you have a great deal of confidence, only some confidence, or hardly any confidence at all in them?) Scientific Community" (1 = a great deal, 2 = only some, 3 = hardly any, D = don't know, I = not applicable, s = skipped on web)

We reverse the measure such that higher values represent higher trust. 

```{r}
# reverse scale and rename
gss_cross <- gss_cross %>%
  mutate(trust = 4 - consci)

# check
gss_cross[gss_cross$year == 1983, c("consci", "trust")]
```

## Covariates

A `year` variable already exists. 

```{r}
gss_cross <- gss_cross %>% 
  mutate(country = "United States", 
         scale = max(gss_cross$trust, na.rm = TRUE)
         )
```

### Other

The GSS includes other institutional trust items that can serve as a baseline to evaluate the evolution of trust in science. 

-   [`CONFINAN`] : "Banks and financial institutions"
-   [`CONBUS`] : "Major companies"
-   [`CONCLERG`] : "Organized religion"
-   [`CONEDUC`] : "Education"
-   [`CONFED`] : "Executive branch of the federal government"
-   [`CONLABOR`] : "Organized Labor"
-   [`CONPRESS`] : "Press"
-   [`CONMEDIC`] : "Medicine"
-   [`CONTV`] : "TV"
-   [`CONJUDGE`] : "U.S. Supreme Court"
-   [`CONSCI`] : "Scientific community"
-   [`CONLEGIS`] : "Congress"
-   [`CONARMY`] : "Military"

All these variables are measured as follows: "I am going to name some institutions in this country. As far as the people running these institutions are concerned, would you say you have a great deal of confidence, only some confidence, or hardly any confidence at all in them?" {each of the 16 names of the above organizations} (1) A great deal confidence, (2) Only some confidence, (3) Hardly any confidence at all"

```{r}
# vector with all institutional trust variable names
institutional_trust_variables <- c("conbus", "conclerg", "coneduc", "conlabor", "conpress", "contv", "conmedic", "confed", "conlegis", "conjudge", "consci",   "conarmy", "confinan") 

# See in which year which measure was collected
gss_cross %>% group_by(year) %>% 
summarize(across(all_of(institutional_trust_variables), ~ mean(.x, na.rm=TRUE)
                 )
          )
```

Just as for `CONSCI` before, we reverse the coding for all institutional trust variables, such that higher values represent higher trust.

```{r}
# mini function that reverses the scale (such that higher = more confidence)
reverse_scale <- function(x) (4-x)

# check
gss_cross %>% mutate(across(all_of(institutional_trust_variables), reverse_scale, .names = "recoded_{col}")) %>% 
  select(recoded_conbus, conbus) %>% 
  # make sure to select valid cases to check (in the first year, 
  # trust in institutions wasn't collected) 
  filter(!is.na(conbus))

# recode
gss_cross <- gss_cross %>%
  mutate(across(all_of(institutional_trust_variables), reverse_scale)) 
```
## Export

```{r}
gss_cross_cleaned <- gss_cross %>% 
  mutate(data = "gss_cross") %>% 
  select(trust, scale, country, year, starts_with("con"), data)

write_csv(gss_cross_cleaned, "data/gss_cross/gss_cross_cleaned.csv")
```


# Wellcome Global Monitor 2018 {wgm_2018}

```{r, message=FALSE}
wgm_2018 <- read_csv("data/wgm_2018/wgm_2018.csv")
```

## Trust

- `Q11C`: "How much do you trust each of the following? How about scientists in this country? Do you trust them a lot, some, not much, or not at all?" (1=A lot, 2=Some, 3=Not much, 4=Not at all, 98=(DK), 99=(Refused))

We re-label `NA`s. According to the codebook, `NA`s are coded as `98` and `99`. 

```{r}
# check response levels
levels(as.factor(wgm_2018$Q11C))
```

```{r}
# define NA values
NA_responses <- c(98, 99)
# mini function to assign NA's to values defined above
recodeNAs <- function(x) (ifelse (x %in% NA_responses, NA, x))

# check
wgm_2018 %>% 
  # make sure answers written to lower case
  mutate(across(Q11C, recodeNAs, .names = "{.col}_recoded")
         ) %>% 
  select(Q11C, Q11C_recoded) %>% 
  filter(is.na(Q11C_recoded))

# recode
wgm_2018 <- wgm_2018 %>% 
  # make sure answers written to lower case
  mutate(across(Q11C, recodeNAs)
         ) 

# check
table(wgm_2018$Q11C, useNA = "always")
```

We reverse the coding such that higher values represent higher trust. 

```{r}
wgm_2018 <- wgm_2018 %>% 
  mutate(trust = 5 - Q11C)

# check
wgm_2018[1:100, c("trust", "Q11C")]
```

## Covariates

```{r}
wgm_2018 <- wgm_2018 %>% 
  mutate(year = 2018,
         scale = max(trust, na.rm = TRUE)
         )
```

### Country

Country codes are available from the codebook.

```{r}
wgm_2018$WP5 <- as.factor(wgm_2018$WP5)
levels(wgm_2018$WP5)

wgm_2018$country <- wgm_2018$WP5 %>%
  dplyr::recode(`1`= "United States", `2` ="Egypt", `3` ="Morocco", `4` ="Lebanon", `5` ="Saudi Arabia", `6` ="Jordan", `8` ="Turkey", `9` ="Pakistan", `10` ="Indonesia", `11` ="Bangladesh", `12` ="United Kingdom", `13` ="France", `14` ="Germany", `15` ="Netherlands", `16` ="Belgium", `17` ="Spain", `18` ="Italy", `19` ="Poland", `20` ="Hungary", `21` ="Czech Republic", `22` ="Romania", `23` ="Sweden", `24` ="Greece", `25` ="Denmark", `26` ="Iran", `28` ="Singapore", `29` ="Japan", `30` ="China", `31` ="India", `32` ="Venezuela", `33` ="Brazil", `34` ="Mexico", `35` ="Nigeria", `36` ="Kenya", `37` ="Tanzania", `38` ="Israel", `39` ="Palestinian Territories", `40` ="Ghana", `41` ="Uganda", `42` ="Benin", `43` ="Madagascar", `44` ="Malawi", `45` ="South Africa", `46` ="Canada", `47` ="Australia", `48` ="Philippines", `49` ="Sri Lanka", `50` ="Vietnam", `51` ="Thailand", `52` ="Cambodia", `53` ="Laos", `54` ="Myanmar", `55` ="New Zealand", `57` ="Botswana", `60` ="Ethiopia", `61` ="Mali", `62` ="Mauritania", `63` ="Mozambique", `64` ="Niger", `65` ="Rwanda", `66` ="Senegal", `67` ="Zambia", `68` ="South Korea", `69` ="Taiwan", `70` ="Afghanistan", `71` ="Belarus", `72` ="Georgia", `73` ="Kazakhstan", `74` ="Kyrgyzstan", `75` ="Moldova", `76` ="Russia", `77` ="Ukraine", `78` ="Burkina Faso", `79` ="Cameroon", `80` ="Sierra Leone", `81` ="Zimbabwe", `82` ="Costa Rica", `83` ="Albania", `84` ="Algeria", `87` ="Argentina", `88` ="Armenia", `89` ="Austria", `90` ="Azerbaijan", `96` ="Bolivia", `97` ="Bosnia and Herzegovina", `99` ="Bulgaria", `100` ="Burundi", `103` ="Chad", `104` ="Chile", `105` ="Colombia", `106` ="Comoros", `108` ="Republic of Congo", `109` ="Croatia", `111` ="Cyprus", `114` ="Dominican Republic", `115` ="Ecuador", `116` ="El Salvador", `119` ="Estonia", `121` ="Finland", `122` ="Gabon", `124` ="Guatemala", `125` ="Guinea", `128` ="Haiti", `129` ="Honduras", `130` ="Iceland", `131` ="Iraq", `132` ="Ireland", `134` ="Ivory Coast", `137` ="Kuwait", `138` ="Latvia", `140` ="Liberia", `141` ="Libya", `143` ="Lithuania", `144` ="Luxembourg", `145` ="Macedonia", `146` ="Malaysia", `148` ="Malta", `150` ="Mauritius", `153` ="Mongolia", `154` ="Montenegro", `155` ="Namibia", `157` ="Nepal", `158` ="Nicaragua", `160` ="Norway", `163` ="Panama", `164` ="Paraguay", `165` ="Peru", `166` ="Portugal", `173` ="Serbia", `175` ="Slovakia", `176` ="Slovenia", `183` ="Eswatini", `184` ="Switzerland", `185` ="Tajikistan", `186` ="The Gambia", `187` ="Togo", `190` ="Tunisia", `191` ="Turkmenistan", `193` ="United Arab Emirates", `194` ="Uruguay", `195` ="Uzbekistan", `197` ="Yemen", `198` ="Kosovo", `202` ="Northern Cyprus", .default = NA_character_)

# check 
wgm_2018 %>% 
  group_by(country, WP5) %>% 
  summarize(count = n())
```

## Export

```{r}
wgm_2018_cleaned <- wgm_2018  %>% 
  mutate(data = "wgm_2018 ") %>% 
  select(trust, scale, country, year, data)

write_csv(wgm_2018_cleaned, "data/wgm_2018/wgm_2018_cleaned.csv")
```

# Wellcome Global Monitor 2020 {wgm_2020}

```{r, message=FALSE}
wgm_2020 <- read_csv("data/wgm_2020/wgm_2020.csv")
```

## Trust

- `W5C`: "How much do you trust each of the following? Do you trust them a lot, some, not much,
or not at all? If you don’t know, please just say so. How about scientists in this country?" (1=A lot, 2=Some, 3=Not much, 4=Not at all, 99=DK/Refused)

We re-label `NA`s. According to the codebook, `NA`s are coded as `98` and `99`. 

```{r}
# check response levels
levels(as.factor(wgm_2020$W5C))
```

```{r}
# define NA values
NA_responses <- c(99)
# mini function to assign NA's to values defined above
recodeNAs <- function(x) (ifelse (x %in% NA_responses, NA, x))

# check
wgm_2020 %>% 
  # make sure answers written to lower case
  mutate(across(W5C, recodeNAs, .names = "{.col}_recoded")
         ) %>% 
  select(W5C, W5C_recoded) %>% 
  filter(is.na(W5C_recoded))

# recode
wgm_2020 <- wgm_2020 %>% 
  # make sure answers written to lower case
  mutate(across(W5C, recodeNAs)
         ) 

# check
table(wgm_2020$W5C, useNA = "always")
```

We reverse the coding such that higher values represent higher trust. 

```{r}
# reverse scale
wgm_2020 <- wgm_2020 %>% 
  mutate(trust = 5 - W5C)

# check
wgm_2020[1:100, c("trust", "W5C")]
```

## Covariates

```{r}
wgm_2020 <- wgm_2020 %>% 
  mutate(year = 2020, 
         country = COUNTRYNEW,
         scale = max(trust, na.rm = TRUE))
```

## Export

```{r}
wgm_2020_cleaned <- wgm_2020  %>% 
  mutate(data = "wgm_2020 ") %>% 
  select(trust, scale, country, year, data)

write_csv(wgm_2020_cleaned, "data/wgm_2020/wgm_2020_cleaned.csv")
```

# KGSS (General Social Survey South Korea) {kgss}

```{r, message=FALSE}
kgss <- read_sav("data/kgss/2003-2021_KGSS_ENG_public_v1_07012022.sav")
```

## Trust

-   `CONSCI`: "I am going to name some institutions in this country. As far as the people running these institutions are concerned, would you say you have a great deal of confidence, only some confidence, or hardly any confidence at all in them?... Scientific community" (1 = a great deal, 2 = only some, 3 = hardly any confidence at all, -8 = don't know, -1 = not applicable)

We re-code `NA`s and reverse the scale such that higher values represent higher trust. 

```{r}
levels(as.factor(kgss$CONSCI))

# Re-code NA's, reverse scale (such that higher = more trust) and rename variable
kgss <- kgss %>% 
  mutate(trust = ifelse(CONSCI < 0, NA, # recode NAs
                              4 - CONSCI)) # reverse scale

# check
kgss[1:100, c("CONSCI", "trust")]
```

## Covariates

```{r}
kgss <- kgss %>% 
  mutate(country = "South Korea", 
         year = YEAR,
         scale = max(trust, na.rm = TRUE)
  )
```

## Export

```{r}
kgss_cleaned <- kgss  %>% 
  mutate(data = "kgss ") %>% 
  select(trust, scale, country, year, data)

write_csv(kgss_cleaned, "data/kgss/kgss_cleaned.csv")
```

# Pew International Science Survey {pew_international}

```{r, message=FALSE}
# read data
pew_international <- read_sav("./data/pew_international/2019-2020 Pew Research Center International Science Survey.sav")
```

## Trust

-   `Q2d`: "Q2d. How much do you trust ____ to do what is right for (survey public) — a lot, some, not too much, or not at all? d. scientists" (1=A lot, 2=Some, 3=Not too much, 4=Not at all, 99=DK/Refused)

We re-label `NA`s. According to the codebook, `NA`s are coded as `98` and `99`. 

```{r}
# check response levels
levels(as.factor(pew_international$Q2d))
```

```{r}
# define NA values
NA_responses <- c(99)
# mini function to assign NA's to values defined above
recodeNAs <- function(x) (ifelse (x %in% NA_responses, NA, x))

# recode
pew_international <- pew_international %>% 
  # make sure answers written to lower case
  mutate(across(Q2d, recodeNAs)
         ) 

# check
table(pew_international$Q2d, useNA = "always")
```

We reverse the coding such that higher values represent higher trust. 

```{r}
pew_international <- pew_international %>% 
  mutate(trust = 5 - Q2d)

# check
wgm_2020[1:100, c("trust", "W5C")]
```

## Covariates

```{r}
pew_international <- pew_international %>% 
  mutate(
    # Values are numeric, but country names are stored in labels. Replace values with labels.
    country = sjlabelled::to_character(pew_international$place),
    # survey undertaken between end 2019, beginning 2020
    year = 2020, 
    scale = max(trust, na.rm = TRUE)
  )
```

## Export

```{r}
pew_international_cleaned <- pew_international  %>% 
  mutate(data = "pew_international ") %>% 
  select(trust, scale, country, year, data)

write_csv(pew_international_cleaned, "data/pew_international/pew_international_cleaned.csv")
```

# Pew American Trends Panel {pew_trends}

This panel study consists of several different data sets. We store them in a list of data frames. 

```{r}
# Set the path to the directory containing subfolders with .sav files
base_path <- "data/pew_trends/"

# Get a list of subfolders
subfolders <- list.dirs(base_path, full.names = FALSE, recursive = FALSE)

# Initialize an empty list to store your data frames
pew_trends <- list()

# Loop through each subfolder
for (subfolder in subfolders) {
  # Get a list of .sav files in the subfolder
  sav_files <- list.files(file.path(base_path, subfolder), pattern = "\\.sav$", full.names = TRUE)

  # Read each .sav file and store it in the list
  pew_trends[subfolder] <- lapply(sav_files, read_sav)
}

# Now, pew_trends contains all data frames as list elements with names corresponding to subfolder names
```

## Trust

### Identify outcome across survey waves

- 2016, May: `CONFD2_W17`

```{r}
sjlabelled::get_label(pew_trends$W17_May16$CONFD2_W17)
```

```{r}
sjlabelled::val_labels(pew_trends$W17_May16$CONFD2_W17)
```

- 2018, Feb: `CONFF_W31`

```{r}
sjlabelled::get_label(pew_trends$W31_Feb18$CONFF_W31)
```

```{r}
sjlabelled::val_labels(pew_trends$W31_Feb18$CONFF_W31)
```

- 2018, Nov: `CONF1f_W40`

```{r}
sjlabelled::get_label(pew_trends$W40_Nov18$CONF1f_W40)
```

```{r}
sjlabelled::val_labels(pew_trends$W40_Nov18$CONF1f_W40)
```

- 2019, Jan: `CONFd_F2_W42`

```{r}
sjlabelled::get_label(pew_trends$W42_Jan19$CONFd_F2_W42)
```

```{r}
sjlabelled::val_labels(pew_trends$W42_Jan19$CONFd_F2_W42)
```

- 2020, Apr: `CONF_g_W66`

```{r}
sjlabelled::get_label(pew_trends$W66_Apr20$CONF_g_W66)
```

```{r}
sjlabelled::val_labels(pew_trends$W66_Apr20$CONF_g_W66)
```

- 2020, Nov: `CONF_g_W79`

```{r}
sjlabelled::get_label(pew_trends$W79_Nov20$CONF_g_W79)
```

```{r}
sjlabelled::val_labels(pew_trends$W79_Nov20$CONF_g_W79)
```

- 2021, Dec: `CONF_g_W100`

```{r}
sjlabelled::get_label(pew_trends$W100_Dec21$CONF_g_W100)
```

```{r}
sjlabelled::val_labels(pew_trends$W100_Dec21$CONF_g_W100)
```

- 2021, Dec: `CONF_g_W114`

```{r}
sjlabelled::get_label(pew_trends$W114_Sept22$CONF_g_W114)
```

```{r}
sjlabelled::val_labels(pew_trends$W114_Sept22$CONF_g_W114)
```

### Rename 

```{r}
# List of original trust variable names
original_trust_names <- c("CONFD2_W17", "CONFF_W31", "CONF1f_W40", "CONFd_F2_W42", "CONF_g_W66", "CONF_g_W79", "CONF_g_W100", "CONF_g_W114")

# Function to rename trust variable in each data frame
rename_trust_variable <- function(df) {
  df %>%
    rename(trust = any_of(original_trust_names))
}

# Apply the function to each data frame in the list
pew_trends <- lapply(pew_trends, rename_trust_variable)

# Now, pew_trends_renamed contains the data frames with the trust variable renamed to "trust"
```

### Rervse and code NA's

We code invalid responses as `NA` and reverse the value coding such that higher values represent higher trust. 

```{r}
# define NA values
NA_responses <- c(99)
# mini function to assign NA's to values defined above
recodeNAs <- function(x) (ifelse (as.numeric(x) %in% NA_responses, NA, x))

# recode NAs in data frame and reverse scale
recode_single_data_frame <- function(df) {
  df <- df %>% 
    # recode NAs
    mutate(across(trust, recodeNAs)) %>% 
    # reverse scale
    mutate(across(trust, ~ 5-.x )) 
}

# Apply the function to each data frame in the list
pew_trends <- lapply(pew_trends, recode_single_data_frame)
```

### Participants across waves

From the codebook, we know that the variable `QKEY` is a unique identifier assigned to each respondent. `QKEY` can be used to link multiple panel waves together.

```{r}
# Create an empty data frame to store participant information
participant_info <- data.frame()

# Loop through each wave in the list
for (wave_name in names(pew_trends)) {
  # Get the data frame for the current wave
  current_wave <- pew_trends[[wave_name]]

  # Extract unique participants (QKEY) in the current wave
  participants_in_wave <- unique(current_wave$QKEY)

  # Create a data frame with participant information for the current wave
  wave_participant_info <- data.frame(QKEY = participants_in_wave, wave = wave_name)

  # Append the participant information to the main data frame
  participant_info <- rbind(participant_info, wave_participant_info)
}

nrow(participant_info)
```

```{r}
# summarize participant data
summary <- participant_info %>% 
  group_by(QKEY) %>% 
  summarise(n_waves = n_distinct(wave))

# as bar plot
ggplot(summary, aes(x = n_waves)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, show.legend = FALSE) +
  labs(x = "Waves per participant",
       y = "N Participants present") +
  theme_minimal()
```

This looks promising. However, in some survey waves, only a randomized subset of the sample was asked the trust in scientists question. 

So, next, we want to know how many participants have valid responses per wave. 

```{r}
# Create an empty data frame to store participant information
participant_info <- data.frame()

# Loop through each wave in the list
for (wave_name in names(pew_trends)) {
  # Get the data frame for the current wave
  current_wave <- pew_trends[[wave_name]]
  
  # filter out NA responses for trust
  current_wave <- current_wave %>% drop_na(trust)

  # Extract unique participants (QKEY) in the current wave
  participants_in_wave <- unique(current_wave$QKEY)

  # Create a data frame with participant information for the current wave
  wave_participant_info <- data.frame(QKEY = participants_in_wave, wave = wave_name)

  # Append the participant information to the main data frame
  participant_info <- rbind(participant_info, wave_participant_info)
}

nrow(participant_info)
```

```{r}
# summarize participant data
summary <- participant_info %>% 
  group_by(QKEY) %>% 
  summarise(n_waves = n_distinct(wave))

# as bar plot
ggplot(summary, aes(x = n_waves)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, show.legend = FALSE) +
  labs(x = "Waves per participant",
       y = "N Participants valid answers") +
  theme_minimal()
```

## Covariates

Before building a common data frame, we will have to assign common names to covariates like we did for the `trust` variable above.

Covariates we want to include are: 

- 

```{r}
# # List of original trust variable names
# original_trust_names <- c("CONFD2_W17", "CONFF_W31", "CONF1f_W40", "CONFd_F2_W42", "CONF_g_W66", "CONF_g_W79", "CONF_g_W100", "CONF_g_W114")
# 
# # Function to rename trust variable in each data frame
# rename_trust_variable <- function(df) {
#   df %>%
#     rename(trust = any_of(original_trust_names))
# }
# 
# # Apply the function to each data frame in the list
# pew_trends <- lapply(pew_trends, rename_trust_variable)
# 
# # Now, pew_trends_renamed contains the data frames with the trust variable renamed to "trust"
```

### Date

```{r}
# Function to extract month and year from the data set name and add a new variable
add_date_year <- function(df, wave_name) {
  # Extract month and year from the name
  month_year <- gsub(".*_(\\D+)(\\d+)", "\\1 \\2", wave_name)
  
  # Create a new variable 'date' with the corresponding date
  df <- df %>% 
    mutate(date = as.Date(paste(month_year, "01", sep = " "), 
                          format = "%B %y %d"),
           year = year(date)
           
           )
  
  return(df)
}

# Apply the function to each data frame in the list
pew_trends <- Map(add_date_year, pew_trends, names(pew_trends))

# Now, pew_trends contains the data frames with a new 'date' and 'year' variable
```

```{r}
# Reduce all data frames in the list to variables starting with "CONF"
pew_trends <- lapply(pew_trends, function(df) {
  select(df, c(QKEY, trust, date, year))
})

# convert the list into a data frame
pew_trends <- bind_rows(pew_trends) %>% 
  rename(id = QKEY)
```

## Covariates

```{r}
pew_trends <- pew_trends %>% 
  mutate(country = "United States", 
         scale = max(trust, na.rm = TRUE)
         )
```

## Export

```{r}
pew_trends_cleaned <- pew_trends  %>% 
  mutate(data = "pew_trends") %>% 
  select(id, trust, scale, date, year, country, data)

write_csv(pew_trends_cleaned, "data/pew_trends/pew_trends_cleaned.csv")
```

# General Social Survey Panel (US) {gss_panel}

The 2016 - 2020 file has only two waves (16 and 18 are two different panels, and 20 is the second wave for these two, see codebook)

```{r}
# Set the path to the directory containing .dta files
data_folder <- "data/gss_panel/"

# List all .dta files in the folder
dta_files <- list.files(data_folder, pattern = "\\.dta$", full.names = TRUE)

# read all .dta files in the folder into a list of data frames
gss_panel <- setNames(lapply(dta_files, read_dta), 
                  tools::file_path_sans_ext(basename(dta_files)))

# change names
names(gss_panel)
new_names <- c("panel_06", "panel_08", "panel_10", "panel_20")
names(gss_panel) <- new_names

```

## Trust

`consci_*`: "I am going to name some institutions in this country. As far as the people running these institutions are concerned, would you say you have
a great deal of confidence, only some confidence, or hardly any confidence at all in them?"

```{r}
sjlabelled::get_label(gss_panel$panel_06$consci_1)
```

```{r}
sjlabelled::val_labels(gss_panel$panel_06$consci_1)
```

## Reduce and reshape data

We start with reducing the data in order to reduce processing time. When reshaping, we also rename the `consci` variable into `trust`. 

```{r}
reshape_data <- function(df, panel_name) {
  
  # reduce data frame to key variables
  reduced <- df %>%
    select(starts_with("id_1"), starts_with("year_"), starts_with("con")) %>%
    # add variables
    mutate(panel = panel_name, 
           country = "United States") %>% 
    # Remove value labels from all variables in the data frame
    mutate(across(everything(), ~haven::zap_labels(.))) %>% 
    # rename id variable
    rename(id = starts_with("id"))
  
  
  # shape to long format
  reshaped <- reduced %>%
    pivot_longer(!c(starts_with("id"), panel, country), names_to = "oldname_wave", values_to = "score") %>%
    separate_wider_delim(oldname_wave, delim = "_", names = c("old_name", "wave")) %>%
    pivot_wider(names_from = old_name, values_from = score) %>% 
    # rename trust variable
    mutate(trust = consci)
  
  return(reshaped)
}

# Apply the function to each data frame in the list
gss_panel <- lapply(names(gss_panel), function(panel_name) {
  reshape_data(gss_panel[[panel_name]], panel_name)
})

names(gss_panel) <- new_names

# The 2020 panel has two id columns, which are assigned to distinct individuals (depending on the starting wave). 
gss_panel$panel_20 %>% 
  filter(!is.na(id1) & !is.na(id2))

# We merge them into one
gss_panel$panel_20 <- gss_panel$panel_20 %>% 
  mutate(id = ifelse(!is.na(id1), paste0(id1, "a"), 
                     paste0(id2, "b"))) %>% 
  select(id, everything()) %>%  # Move 'id' to the first column 
  select(-c(id1, id2))

```

## Rervse and code NA's

We code invalid responses as `NA` and reverse the value coding such that higher values represent higher trust. 

```{r}
# reverse scale
reverse_scales <- function(df) {
  df <- df %>% 
    mutate(across(!c(starts_with("id"), year, panel, wave, country),
                  ~ 4-.x ), 
           id = as.character(id)) 
}

# Apply the function to each data frame in the list
gss_panel <- lapply(gss_panel, reverse_scales)

gss_panel$panel_20
```

## Make common data frame

We store all panels in a single data frame. Importantly, we need to make sure that we create unique `id` values.

```{r}
# turn all into one single data set
gss_panel <- bind_rows(gss_panel) %>% 
  # make a unique id
  mutate(id = paste0(id, "_", panel))
```

### Covariates

We add a scale variable. 

```{r}
gss_panel <- gss_panel %>% 
  mutate(scale = max(trust, na.rm = TRUE))
```


## Export

```{r}
gss_panel_cleaned <- gss_panel  %>% 
  mutate(data = "gss_panel") %>% 
  select(id, trust, scale, year, country, data)

write_csv(gss_panel_cleaned, "data/gss_panel/gss_panel_cleaned.csv")
```

# Algan data

```{r}
algan <- read_dta("data/algan/CAUCP_data_v2.dta")
```

### Trust

```{r}
sjlabelled::get_label(algan$B4_7)
```

```{r}
sjlabelled::val_labels(algan$B4_7)
```

The trust variable is already coded such that higher values represent higher trust, so no need to reverse code. We just rename it below. 

## Covariates

```{r}
# Set options to avoid scientific notation for the id variable and display more digits
options(scipen = 999, digits = 12)
```

```{r}
algan <- algan %>% 
  # rename and add covariates
  mutate(
    date = as.Date(DATEINTER, format = "%Y%m%d"),
    year = year(date), 
    trust = B4_7,
    country = sub(".*\\.", "", sjlabelled::to_character(algan$country)), 
    scale = max(trust, na.rm = TRUE)
  ) %>% 
  # reduce data
  select(country, id, wave, date, trust, year, scale)

algan  
```

## Export

```{r}
algan_cleaned <- algan  %>% 
  mutate(data = "algan") %>% 
  select(id, trust, scale, date, year, country, data)

write_csv(algan_cleaned, "data/algan/algan_cleaned.csv")
```

# Panel data Daniel 

```{r, message=FALSE}
nettle <- read_csv("data/nettle/CCL_processed_data.csv")
```

## Trust

`sciencetrust`: On a scale from 1-100, to what extent do you trust respectively: ...scientists (measured from February 2023 onwards)

## Covariates

### Date

```{r}
# check dates and transformation
nettle$month_as_date <- as.Date(paste(nettle$month, "01", sep = "-"), 
                          format = "%y-%b-%d")

nettle$survey_date <- as.Date(
  as.POSIXct(nettle$StartDate, format = "%m/%d/%Y %H:%M", tz = "UTC")
)

nettle %>% select(month_as_date , month, StartDate, survey_date)
```

The transformation is coherent with the old `month` variable, but the `StartDate` variable is sometimes later. From the codebook, we know that we should rely on the `StartDate` variable (the `month` variable is important for financial questions, which refer to the month preceeding the survey wave). 

```{r}
nettle <- nettle %>% 
  # rename and add covariates
  mutate(
    date = as.Date(
      as.POSIXct(StartDate, format = "%m/%d/%Y %H:%M", tz = "UTC")
    ),
    year = year(date), 
    trust = sciencetrust,
    scale = max(trust, na.rm = TRUE), 
    id = pid
  ) %>% 
  # reduce data
  select(country, id, date, trust, year, scale)

nettle
```
## Export

```{r}
nettle_cleaned <- nettle  %>% 
  mutate(data = "nettle") %>% 
  select(id, trust, scale, date, year, country, data)

write_csv(nettle_cleaned, "data/nettle/nettle_cleaned.csv")
```

# TISP data (for now only for France)


# German Wissenschaftsbarometer

```{r, message=FALSE}
barometer <- read_sav("data/wissenschaftsbarometer/Datensatz_gemergt_final.sav")
```

## Trust

The data contains several trust questions.

```{r}
# select all trust-related variables
trust <- barometer %>% select(starts_with("trust"))

# Extract variable names, labels, and value labels
trust_items <- names(trust) %>%
  map_df(~ {
    var_name <- .x
    var_label <- attr(barometer[[.x]], "label")
    value_labels <- attr(barometer[[.x]], "labels")
    
    # Combine value labels into a single string
    value_labels_str <- paste(names(value_labels), value_labels, sep = "=", collapse = "; ")
    
    tibble(Name = var_name, 
           Label = var_label, 
           Value_Labels = value_labels_str)
  })

# Print the data frame
trust_items
```

```{r}
# Get an overview of labels for the trust-related variables
value_labels <- trust %>%
  map(~ {
    labels <- attr(., "labels")
    tibble(value = names(labels), label = labels)
  }) %>%
  bind_rows(.id = "Variable")

# check if they are always the same
value_labels %>% 
  group_by(label) %>% 
  reframe(value = unique(value))
# yes

```

Were all items asked in all waves? 

```{r}
barometer %>% 
  pivot_longer(starts_with("trust"), 
               names_to = "trust_item",
               values_to = "score") %>% 
  group_by(trust_item) %>% 
  drop_na(score) %>% 
  summarize(waves = n_distinct(wave), 
            subjects = n_distinct(id))
```

The main variable that has been asked in 9 survey waves is:

`trust1`: "Wie sehr vertrauen Sie Wissenschaft und Forschung?"

```{r}
table(barometer$trust1)
```

### Reverse and code NA's

```{r}
# Function to recode and reverse scale
recode_and_reverse <- function(data, vars) {
  data %>%
    # make "6" NA
    mutate(across(all_of(vars), ~ ifelse(. == 6, NA, .)
                  )
           ) %>%
    # reverse values for 1 to 5
    mutate(across(all_of(vars), ~ ifelse(!is.na(.), 6 - ., .)
                  )
           )
}


# Specify the variables to recode and reverse
trust_items <- barometer %>% select(starts_with("trust")) %>% names(.)

# Apply the function to the specified variables
barometer <- recode_and_reverse(barometer, trust_items) 
```

## Covariates

### Date

```{r}
# Function to convert wave values to dates and years
convert_wave_to_date_and_year <- function(data) {
  
  # Extract the wave labels
  wave_labels <- attr(data$wave, "labels")
  
  # Create a named vector to map wave values to their labels
  wave_to_date <- setNames(names(wave_labels), wave_labels)
  
  
  data %>%
    mutate(
      wave_label = as.character(wave_to_date[as.character(wave)]),
      date = as.Date(paste0(wave_label, "/01"), format = "%Y/%m/%d"),
      year = year(date)
    )
}

# Apply the function to the barometer dataset
barometer <- convert_wave_to_date_and_year(barometer)

# test
# test <- convert_wave_to_date_and_year(barometer)
# table(test$year)
```

```{r}
barometer %>% group_by(id) %>% 
  summarise(date = n_distinct(date)) %>% 
  group_by(date) %>% 
  summarise(subjects = n_distinct(id))
```

```{r}
barometer <- barometer %>% 
  # rename and add covariates
  mutate(
    date = date,
    year = year, 
    trust = trust1,
    scale = max(trust, na.rm = TRUE), 
    id = id, 
    country = "Germany"
  ) %>% 
  # reduce data
  select(country, id, date, trust, year, scale)
```
## Export

```{r}
barometer_cleaned <- barometer  %>% 
  mutate(data = "Wissenschaftsbarometer") %>% 
  select(id, trust, scale, date, year, country, data)

write_csv(barometer_cleaned, "data/wissenschaftsbarometer/wissenschaftsbarometer_cleaned.csv")
```

## Some example change

Hi Jan !

blablabla
ofpjfeofe

This is good, but I would add this. 

```{r}
#hi
7+3
```

petit test

