---
title: "descriptives"
output: html_document
date: "2022-12-15"
bibliography: references.bib
---

```{r, message=FALSE}
library(tidyverse)
library(haven) # read stata data
library(patchwork) # for combining plots
```

## Load data

```{r, message=FALSE}
# Set the path to the directory containing data files
data_folder <- "data/"

# List all files in the folder
file_names <- list.files(data_folder, pattern = "cleaned.csv", recursive = TRUE,
                        full.names = TRUE)

# read all files in the folder into a list of data frames
data_list <- setNames(lapply(file_names, read_csv), 
                  sub("_cleaned", "", tools::file_path_sans_ext(basename(file_names))))

# check names
names(data_list)
```

## Make common data frame

```{r}
# check id column 
# (necessary to be the same data type when merging to a single data frame)
summarize_id_data_types <- function(data_list) {
  # Create an empty data frame to store the summary
  summary_df <- data.frame(DataFrame = character(), DataType = character(), stringsAsFactors = FALSE)
  
  # Loop through each data frame in the list
  for (i in seq_along(data_list)) {
    df <- data_list[[i]]
    
    # Check if 'id' column exists in the data frame
    if ("id" %in% colnames(df)) {
      # Get data type of 'id' column
      data_type <- typeof(df$id)
      
      # Add the summary to the data frame
      summary_df <- rbind(summary_df, data.frame(DataFrame = names(data_list)[i], DataType = data_type, stringsAsFactors = FALSE))
    }
  }
  
  return(summary_df)
}

# Assuming data_list is your list of data frames
summary_result <- summarize_id_data_types(data_list)

# View the summary
print(summary_result)
```


```{r}
# Define a function to turn the id variable into character for all data sets
id_to_character <- function(df) {
  
  # Check if 'id' column exists in the data frame
  if ("id" %in% colnames(df)) {
    df %>%
      mutate(id = as.character(id)) 
  } else {
      return(df)
    }
}

# Use lapply to apply the function to each data frame in the list
data_list <- lapply(data_list, id_to_character)

# Combine the individual data frames into a single data frame
data <- bind_rows(data_list, .id = "file_name")
```

## Overview of time span

```{r}
overview_time_span <- data %>% 
  group_by(data) %>% 
  reframe(year = unique(year)) %>% 
  drop_na(year)

ggplot(data = overview_time_span,
       aes(x = year, y = data)) +
  geom_tile() +
  # make more detailed scale
  scale_x_continuous(breaks = seq(1970, 
                                  max(overview_time_span$year), by = 5)) +
  # Add nice labels
  labs(x = "Year", y = NULL) +
  # Force all the tiles to have equal widths and heights
  coord_equal() +
  # Use a cleaner theme
  theme_minimal()
```

## Panel data

# Make a version with acutal data (year/month) on x axis. 

First, we restrict the data to panel data and to valid responses only. 

```{r}
panel_data <- data %>% 
  filter(!is.na(id) & data != "Wissenschaftsbarometer") %>% 
  drop_na(trust)
```

All panel data is in long format. If we group by participant and count the observations, we get the number of waves a participant has been present in. 

```{r}
# count waves per participant
waves_per_participant <- panel_data %>% 
  group_by(id, data) %>% 
  summarize(waves = n())

# as bar plot
ggplot(waves_per_participant, aes(x = waves)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, show.legend = FALSE, 
            size = 2) +
  labs(x = "Waves per participant",
       y = "N Participants (valid answers)") +
  theme_minimal() + 
  facet_wrap(~data)
```
```{r}
# # make one plot for each data frame
# plot_list <- setNames(
#   lapply(names(data_list), function(df_name) {
#     
#     # pick data frame
#     df <- data_list[[df_name]] %>% 
#       drop_na(trust)
#     
#     # Apply everything only to panel data
#     if ("id" %in% colnames(df)) {
#     
#     # count waves per participant
#     waves_per_participant <- df %>% 
#       group_by(id, data) %>% 
#       summarize(waves = n())
#     
#     # as bar plot
#     wave_plot <- ggplot(waves_per_participant, aes(x = waves)) +
#       geom_bar(stat = "count") +
#       geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, show.legend = FALSE) +
#       labs(x = "Waves per participant",
#            y = "N Participants (valid answers)") +
#       theme_minimal()
#     
#     return(wave_plot) 
#     }
#   }),
#   names(data_list)
# )
# 
# plot_list
```

```{r}
# calculate general waves
overview_panel <- panel_data %>% 
  mutate(month = lubridate::round_date(date, unit = "month"))

ggplot(data = overview_panel,
       aes(x = month, y = data)) +
  geom_tile() +
  # make more detailed scale
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y", expand = c(0, 0)) +
  # Add nice labels
  labs(x = "Date", y = NULL) +
  # Force all the tiles to have equal widths and heights
  coord_equal() +
  # Use a cleaner theme
  theme_minimal() 
```


One plot for each panel: 

check only valid responses, 

make a plot as for pew (left)

group by wave, then plot survey data rounded by month (for algan, nettle and pew, by year for gss), 

```{r}
# summarize participant data
# summary <- participant_info %>% 
#   group_by(QKEY) %>% 
#   summarise(n_waves = n_distinct(wave))

# as bar plot
# ggplot(summary, aes(x = n_waves)) +
#   geom_bar(stat = "count") +
#   geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, show.legend = FALSE) +
#   labs(x = "Waves per participant",
#        y = "N Participants valid answers") +
#   theme_minimal()
```


